---
title: "rent_index_a2020_lag_train_test_all_income"
output: html_notebook
---

```{r}
install.packages("mondate")
library(mondate)

library(stringr)

library(lubridate)

if (!require(remotes)) install.packages("remotes")
remotes::install_github("keithabailey/PRMTools")
library(PRMTools)

if (!require(dplyr)) install.packages("dplyr")
library(dplyr)

suppressPackageStartupMessages(if (!require(xts)) install.packages("xts"))
library(xts)
suppressPackageStartupMessages(if (!require(forecast)) install.packages("forecast") )
library(forecast)

if (!require(rlist)) install.packages("rlist")
library(rlist)

if (!require(Ecdat)) install.packages("Ecdat")
library(Ecdat)

if (!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

if (!require(Hmisc)) install.packages("Hmisc")
library(Hmisc)

if (!require(ISLR)) install.packages("ISLR")
library(ISLR)

if (!require(olsrr)) install.packages("olsrr")
library(olsrr)

if (!require(car)) install.packages("car")
library(car)

if (!require(fastDummies)) install.packages("fastDummies")
library(fastDummies)
```

```{r}
#Read files

rent_data_mod_a2020 <- read.csv("rent_data_mod_a2020.csv")

head(rent_data_mod_a2020)

```

```{r}
q1 = summary(rent_data_mod_a2020$comp_ratio_1000_total)[2]
q3 = summary(rent_data_mod_a2020$comp_ratio_1000_total)[5]
t_min = q1 - (3*IQR(rent_data_mod_a2020$comp_ratio_1000_total))
t_max = q3 + (3*IQR(rent_data_mod_a2020$comp_ratio_1000_total))
rent_data_mod_a2020 <- rent_data_mod_a2020%>% mutate(outlier = case_when(comp_ratio_1000_total<t_min ~ -1, comp_ratio_1000_total>t_max ~ 1, TRUE ~ 0))

```


```{r}
rent_data_mod_a2020$month <- format(as.Date(rent_data_mod_a2020$zillow_month_day),"%m")
nlag = 4
rent_data_mod_a2020_lag <- rent_data_mod_a2020%>%group_by(zillow_zip)%>%arrange(desc(zillow_month_day))%>%mutate(zillow_rent_value_index_lag = lag(zillow_rent_value_index, n = nlag),zillow_month_day_lag = lag(zillow_month_day, n = nlag))


rent_data_mod_a2020_lag <- rent_data_mod_a2020_lag%>%mutate(zillow_month_day_lag = ifelse(!is.na(zillow_rent_value_index_lag), zillow_month_day_lag, format(ceiling_date(as.Date(zillow_month_day), "month") + months(nlag) - days(1),"%Y-%m-%d") ))%>%arrange(zillow_zip,zillow_month_day)

tail(rent_data_mod_a2020_lag%>%filter(zillow_zip == 10001)%>%select(zillow_month_day,zillow_month_day_lag,zillow_rent_value_index,zillow_rent_value_index_lag, outlier),nlag*2)


```

Sales all income levels Train model (train inference window 2021-05-31 - 2022-09-30):

```{r}
rent_data_mod_a2020_lag_train <- rent_data_mod_a2020_lag%>%filter(as.Date(zillow_month_day_lag)<as.Date("2022-10-31"))
tail(rent_data_mod_a2020_lag_train%>%filter(zillow_zip == 10001)%>%select(zillow_month_day,zillow_month_day_lag,zillow_rent_value_index,zillow_rent_value_index_lag, outlier),nlag*2)
list_columns <- colnames(rent_data_mod_a2020_lag_train)
filter_columns = list_columns[grepl("^comp_ratio*|^zillow_rent_value_index|var_zip|^month|^Total_income_per_capita", list_columns)]
filter_columns = filter_columns[!filter_columns %in% c("month_number","zillow_rent_value_index")]
df_lma2020_lag_train <- rent_data_mod_a2020_lag_train%>%filter(outlier==0)%>%filter(!is.na(zillow_rent_value_index_lag))%>%select(filter_columns)
lma2020_lag_train <- lm(zillow_rent_value_index_lag ~ .-zillow_zip, data=df_lma2020_lag_train)  

cat("\nAdjusted R square:")
summary(lma2020_lag_train)$adj.r.squared
cat("\nFStatistic:")
summary(lma2020_lag_train)$fstatistic

cat("\nNumber of significant variables:")
nrow(data.frame(summary(lma2020_lag_train)$coefficients)%>%filter(.[[4]] <0.05))
cat("\nLargest coefficients:")
data.frame(summary(lma2020_lag_train)$coefficients)%>%filter(.[[4]] <0.05)%>% arrange(desc(abs(Estimate)))
sel_columns_lma2020_lag_train <- row.names(data.frame(summary(lma2020_lag_train)$coefficients)%>%filter(.[[4]] <0.05))


```
Optimized train model and test window:

```{r}
df_lma2020_lag_opt_train <- rent_data_mod_a2020_lag_train%>%filter(outlier==0)%>%filter(!is.na(zillow_rent_value_index_lag))%>%select(c(sel_columns_lma2020_lag_train[!grepl("var_zip|^month|*Intercept*",sel_columns_lma2020_lag_train)],var_zip,month,zillow_rent_value_index_lag,zillow_month_day_lag,Total_income_per_capita))

lma2020_lag_opt_train <- lm(zillow_rent_value_index_lag ~ . -zillow_zip, data=df_lma2020_lag_opt_train%>%select(!zillow_month_day_lag))

cat("\nAdjusted R square in train window:")
summary(lma2020_lag_opt_train)$adj.r.squared
cat("\nFStatistic in train window:")
summary(lma2020_lag_opt_train)$fstatistic

cat("\nNumber of significant variables:")
nrow(data.frame(summary(lma2020_lag_opt_train)$coefficients))
cat("\nLargest coefficients:")
data.frame(summary(lma2020_lag_opt_train)$coefficients)%>% arrange(desc(abs(Estimate)))
cat("\nComplaint coefficients:")
data.frame(summary(lma2020_lag_opt_train)$coefficients)[sel_columns_lma2020_lag_train[grepl("comp_ratio|Intercept",sel_columns_lma2020_lag_train)],]%>% arrange(desc(abs(Estimate)))

dfpred_lma2020_lag_opt_test <- rent_data_mod_a2020_lag%>%filter(as.Date(zillow_month_day_lag)>=as.Date("2022-10-31"))%>%filter(outlier==0)%>%filter(!var_zip %in% c("_11249","_11430"))%>%select(zillow_month_day, zillow_rent_value_index,Total_income_per_capita, c(sel_columns_lma2020_lag_train[!grepl("var_zip|^month|*Intercept*",sel_columns_lma2020_lag_train)],var_zip,month,zillow_rent_value_index_lag,zillow_month_day_lag))

dfpred_lma2020_lag_opt_train <- rent_data_mod_a2020_lag%>%filter(as.Date(zillow_month_day_lag)<as.Date("2022-10-31"))%>%filter(outlier==0)%>%filter(!var_zip %in% c("_11249","_11430"))%>%select(c(sel_columns_lma2020_lag_train[!grepl("var_zip|^month|*Intercept*",sel_columns_lma2020_lag_train)],var_zip,month,zillow_rent_value_index_lag,zillow_month_day_lag,Total_income_per_capita))

tail(dfpred_lma2020_lag_opt_test%>%filter(zillow_zip == 10030))

dfpred_lma2020_lag_opt_train$prediction <- predict(lma2020_lag_opt_train, newdata=dfpred_lma2020_lag_opt_train%>%select(!zillow_month_day_lag))

dfpred_lma2020_lag_opt_test$prediction <- predict(lma2020_lag_opt_train, newdata=dfpred_lma2020_lag_opt_test%>%select(!zillow_month_day_lag))

dfpred_lma2020_lag_opt_test$train_prediction <- NA
dfpred_lma2020_lag_opt_train$train_prediction <- dfpred_lma2020_lag_opt_train$prediction
dfpred_lma2020_lag_opt_train$prediction <- NA
dfpred_lma2020_lag_opt_test <- rbind(dfpred_lma2020_lag_opt_train,dfpred_lma2020_lag_opt_test)


y_actual = dfpred_lma2020_lag_opt_test%>%filter(as.Date(zillow_month_day_lag)>=as.Date("2022-10-31")&as.Date(zillow_month_day_lag)<as.Date("2023-02-28"))%>%select(zillow_rent_value_index_lag)
y_pred = dfpred_lma2020_lag_opt_test%>%filter(as.Date(zillow_month_day_lag)>=as.Date("2022-10-31")&as.Date(zillow_month_day_lag)<as.Date("2023-02-28"))%>%select(prediction)

y_actual=y_actual$zillow_rent_value_index_lag
y_pred = y_pred$prediction

cat("\nMean absolute error in prediction window:")
mean(abs((y_actual-y_pred)/y_actual))
cat("\nAdjusted R2 in prediction window:")
k = nrow(data.frame(summary(lma2020_lag_opt_train)$coefficients))
n = length(y_actual)
1 - ((1-cor(y_actual,y_pred)^2)*(n-1)/(n-k-1))



ggplot(data = dfpred_lma2020_lag_opt_test%>%select(zillow_month_day_lag, prediction, train_prediction, zillow_rent_value_index_lag),aes(x = as.Date(zillow_month_day_lag))) + 
  stat_summary(aes(y = prediction, colour = "Predicted Index in test window (n+4)"),geom = "line", fun = "mean", linewidth= 1) +
  stat_summary(aes(y = prediction),geom = "ribbon", fun.data = "mean_cl_normal", alpha = 0.1,fill = "red") +
  stat_summary(aes(y=train_prediction, colour = "Predicted Index in training window"), geom = "line", fun = "mean", linewidth= 1) +
  stat_summary(aes(y = train_prediction),geom = "ribbon", fun.data = "mean_cl_normal", alpha = 0.05,fill = "black") +
  stat_summary(aes(y=zillow_rent_value_index_lag, colour = "Actual Index"), geom = "line", fun = "mean", size= 1) +
  stat_summary(aes(y = zillow_rent_value_index_lag),geom = "ribbon", fun.data = "mean_cl_normal", alpha = 0.1,fill = "blue") +
  scale_colour_manual("", 
                      breaks = c("Predicted Index in test window (n+4)", "Actual Index", "Predicted Index in training window"),
                      values = c("red", "blue", "black")) +
  xlab("Date")+
  ylab("Rent Zillow Home Value Index")+
  ggtitle("2021-2023 Zillow Home Value Time Series \n Forecast Prediction (n+4) vs Actual Index")

```
Identify regression coefficients with NA values and exclude them as it could indicate that variables are linearly related to other variables.
```{r}
# (i.e., lma2020_lag_train)
coef <- data.frame(coef(lma2020_lag_train))
coef$rowname <- rownames(coef)
coef <- coef %>% 
  filter(rowname != "var_zip_11697")

coef_filtered_na <- coef %>% 
  filter(is.na(coef.lma2020_lag_train.))
na_list <- coef_filtered_na$rowname

df_lma2020_lag_train_2 <- df_lma2020_lag_train %>% 
  select(-na_list)

lma2020_lag_train_2 <- lm(zillow_rent_value_index_lag ~ .-
                            zillow_zip -
                            Total_income_per_capita, data = df_lma2020_lag_train_2)
options(max.print=1000000)
summary(lma2020_lag_train_2)
```
Identify regression coefficients with NA values and exclude them as it could indicate that variables are linearly related to other variables.
```{r}
# (i.e., lma2020_lag_opt_train)
coef2 <- data.frame(coef(lma2020_lag_opt_train))
coef2$rowname <- rownames(coef2)
coef2 <- coef2 %>% 
  filter(rowname != "var_zip_11697")

coef_filtered_na2 <- coef2 %>% 
  filter(is.na(coef.lma2020_lag_opt_train.))
na_list <- coef_filtered_na2$rowname

df_lma2020_lag_opt_train_2 <- df_lma2020_lag_opt_train %>% 
  select(-na_list)

lma2020_lag_opt_train_2 <- lm(zillow_rent_value_index_lag ~ .-
                            zillow_zip -
                            Total_income_per_capita, data = df_lma2020_lag_opt_train_2%>%select(!zillow_month_day_lag))
options(max.print=1000000)
summary(lma2020_lag_opt_train_2)
```


Perform variable selection using two different approaches 1) forward stepwise selection based on p-values and 2) backward elimination based on p-values.
```{r}
# Variable selection for the unoptimized model (i.e., lma2020_lag_train)
# forward selection
v_select1 <- ols_step_forward_p(lma2020_lag_train_2)
v_select1
plot(v_select1)

# backward selection
v_select2 <- ols_step_backward_p(lma2020_lag_train_2)
v_select2
plot(v_select2)

# Variable selection for the optimized model (i.e., lma2020_lag_opt_train)
# forward selection
v_select3 <- ols_step_forward_p(lma2020_lag_opt_train_2)
v_select3
plot(v_select3)

# backward selection
v_select4 <- ols_step_backward_p(lma2020_lag_opt_train_2)
v_select4
plot(v_select4)
# vif(lma2020_lag_train_2)

# ols_coll_diag(lma2020_lag_train_2)
```



```{r}
# Extract the coefficients from the summary
coef_names <- row.names(summary(lma2020_lag_train_2)$coefficients)

# Count the number of coefficients that contain "var_zip" in their names
num_var_zip_coef <- sum(grepl("var_zip[_0-9]*", coef_names))
num_var_zip_coef

# Count the number of coefficients that contain "comp_" in their names
num_comp_coef <- sum(grepl("comp_[[:alnum:]]+", coef_names))
num_comp_coef

# count the number of coefficients in the model.
length(coef(lma2020_lag_train_2))

# Extract the coefficients from the summary
coef_names2 <- row.names(summary(lma2020_lag_opt_train_2)$coefficients)

# Count the number of coefficients that contain "var_zip" in their names
num_var_zip_coef2 <- sum(grepl("var_zip[_0-9]*", coef_names2))
num_var_zip_coef2

# Count the number of coefficients that contain "comp_" in their names
num_comp_coef2 <- sum(grepl("comp_[[:alnum:]]+", coef_names2))
num_comp_coef2

# count the number of coefficients in the model.
length(coef(lma2020_lag_opt_train_2))
```
